---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hacksig

<!-- badges: start -->
<!-- badges: end -->

The goal of hacksig is to ...

## Installation

```{r, echo=FALSE, eval=FALSE}
# You can install the released version of hacksig from [CRAN](https://CRAN.R-project.org) with:
# 
# install.packages("hacksig")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Acare/hacksig")
```

## Getting started

The first thing one should do before computing signature scores is to check how 
much of the genes composing a signature are present in your gene expression matrix.
To accomplish this, we can use `check_sig()`:

```{r example, eval=TRUE}
library(hacksig)
check_sig(test_expr)
```

By default, `check_sig()` will compute statistics for every signature implemented 
in `hacksig`. You can filter for specific signatures by entering a keyword in the 
`signatures` argument:

```{r}
check_sig(test_expr, signatures = "immune")
```

The set of available keywords associated to gene signatures can be retrieved with:

```{r}
get_sig_keywords()
```

