---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hacksig

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/hacksig)](https://CRAN.R-project.org/package=hacksig)
[![Codecov test coverage](https://codecov.io/gh/Acare/hacksig/branch/master/graph/badge.svg)](https://app.codecov.io/gh/Acare/hacksig?branch=master)
<!-- badges: end -->

The goal of `hacksig` is to provide a simple and tidy interface to compute gene 
expression signature scores in a number of ways. Several gene signatures and 
methods are implemented, including:

  * [CINSARC](https://doi.org/10.1038/nm.2174)
  * [ESTIMATE](https://doi.org/10.1038/ncomms3612)
  * [Immunophenoscore](https://doi.org/10.1016/j.celrep.2016.12.019)
  * [Cytolitic activity](https://doi.org/10.1016/j.cell.2014.12.033)
  * and many more (see `?signatures_data` for a complete list of gene signatures implemented)

## Installation

```{r, echo=FALSE, eval=FALSE}
# You can install the released version of hacksig from [CRAN](https://CRAN.R-project.org) with:
# 
# install.packages("hacksig")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Acare/hacksig")
```

## Usage

To show example usage of the package, we will use a simulated gene expression matrix `test_expr`, which is stored as an R object in `hacksig`.

```{r, message=FALSE}
library(hacksig)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(future)
library(msigdbr)
```

### Check your signatures

The first thing you should do before computing signature scores is to check how 
much of the genes composing a signature are present in your gene expression matrix.
To accomplish this, we can use `check_sig()`:

```{r}
check_sig(test_expr)
```

By default, `check_sig()` will compute statistics for every signature implemented 
in `hacksig`. You can filter for specific signatures by entering a keyword in the 
`signatures` argument:

```{r}
check_sig(test_expr, signatures = "immune")
```

The set of available keywords associated to gene signatures can be retrieved with:

```{r}
get_sig_keywords()
```

We can also check for signatures not implemented in hacksig, that is custom signatures.
For example, we can use the `msigdbr` package to download the *Hallmark* gene set collection.

```{r}
hallmark_list <- msigdbr(species = "Homo sapiens", category = "H") %>%
    distinct(gs_name, gene_symbol) %>%
    nest(genes = c(gene_symbol)) %>%
    mutate(genes = map(genes, compose(as_vector, unname))) %>%
    deframe()
check_sig(test_expr, hallmark_list)
```

### Compute single sample scores

The main function of the package is `hack_sig()` and it permits to obtain single 
sample scores from gene signatures.
By default, `hack_sig()` will compute scores for all the signatures implemented 
in the package with the original publication method (e.g. weighted sum of expression 
values and model coefficients).

```{r}
hack_sig(test_expr)
```

You can also filter for specific signatures (e.g. the immune and stromal ESTIMATE 
signatures) and choose a particular single sample method:

```{r}
hack_sig(test_expr, signatures = "estimate", method = "zscore")
```

Valid choices for single sample `method`s are:

  * `zscore`, for the combined z-score;
  * `ssgsea`, for the single sample GSEA;
  * `singscore`, for the singscore method.
  
Run `?hack_sig` to see references for these methods.

As in `check_sig()`, the argument `signatures` can also be a list of gene signatures.
For example, we can compute normalized single sample GSEA scores for the Hallmark gene sets:

```{r}
hack_sig(test_expr, hallmark_list, 
         method = "ssgsea", sample_norm = "separate", alpha = 0.5)
```

There are three methods for which `hack_sig()` cannot be used to compute gene signature
scores with the original method. These are: CINSARC, ESTIMATE and the Immunophenoscore.

For the CINSARC classification, you must provide a vector with distant metastasis status:

```{r}
rand_dm <- sample(c(0, 1), size = ncol(test_expr), replace = TRUE)
hack_cinsarc(test_expr, rand_dm)
```

Immune, stromal, ESTIMATE and tumor purity scores from the ESTIMATE method can be obtained with: 

```{r}
hack_estimate(test_expr)
```

Finally, the raw immunophenoscore and its discrete counterpart can be obtained with:

```{r}
hack_immunophenoscore(test_expr)
```


### Classify your samples

If you want to categorize your samples into two or more signature classes based 
on a score cutoff, you can use `hack_class()` after `hack_sig()`:

```{r}
test_expr %>% 
    hack_sig("estimate", method = "singscore", direction = "up") %>% 
    hack_class()
```

### Speed-up computation time

`hacksig` supports the `future` framework for parallel computation. Our single sample 
method implementations based on ranks (i.e. single sample GSEA and singscore) are 
slower than their counterparts implemented in `GSVA` and `singscore`.
Hence, to speed-up computation time you can use the `future` package:

```{r}
plan(multisession)
hack_sig(test_expr, hallmark_list, method = "ssgsea")
```


## Contributing

If you have any suggestions about adding new features to `hacksig`, please 
open an issue request on [GitHub](https://github.com/Acare/hacksig/issues).
Gene-level information about gene signatures are stored in the R object 
`signatures_data` and can be used as a template for requests.

